---
title: "Foehn winds in Switzerland"
author: "Tino Schneidewind"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data, echo=FALSE, message=FALSE}
rm(list= ls())
library(dplyr); library(tidyr); library(ggplot2) ; library(lubridate); library(plotly); 
library(viridis); library(gridExtra); library(corrplot); library(zoo)

# load data
data = read.csv("/Volumes/FS/_ISPM/CCH/Tino/master_thesis/data/Medstat_hospitalizations_aggregated/hosp_buffer_8000.csv") |>
  mutate(date = as.Date(date),
         daymonth = format(date, "%m.%d"))

# seasonal cycle of foehn wind
foehn_d_sum <- data |>
  group_by(date) |>
  summarize(across(f_id, ~ mean(.x, na.rm = TRUE)))

# monthly average of all year-monthly averages
foehn_m_mean <- foehn_d_sum |>
  mutate(year_month =  paste0(format(date, "%Y-%m"),"-01" )) |>
  group_by(year_month) |>
  
  summarize(across(f_id, ~mean(.x, na.rm = TRUE))) |>
  mutate(month =  as.numeric(format(as.Date(year_month), "%m"))) |>
  
  group_by(month) |>
  summarize(across(f_id, ~mean(.x, na.rm = TRUE))) 


# yearly average
foehn_y_sum <- foehn_d_sum |>
  mutate(year =  as.numeric(format(date, "%Y"))) |>
  group_by(year) |>
  summarize(across(f_id, ~mean(.x, na.rm = TRUE)))

# summer average
foehn_y_sum_summer <- foehn_d_sum |>
  mutate(summer =  as.numeric(format(date, "%m"))) |>
  filter(summer >5 & summer < 9) |>
  mutate(year =  as.numeric(format(date, "%Y"))) |>
  group_by(year) |>
  summarize(across(f_id, ~mean(.x, na.rm = TRUE)))

# winter average
foehn_y_sum_winter <- foehn_d_sum |>
  mutate(winter =  as.numeric(format(date, "%m"))) |>
  filter(winter < 3 | winter > 11) |>
  mutate(year =  as.numeric(format(date, "%Y"))) |>
  group_by(year) |>
  summarize(across(f_id, ~mean(.x, na.rm = TRUE)))

# spring average
foehn_y_sum_spring <- foehn_d_sum |>
  mutate(winter =  as.numeric(format(date, "%m"))) |>
  filter(winter > 2 & winter < 6) |>
  mutate(year =  as.numeric(format(date, "%Y"))) |>
  group_by(year) |>
  summarize(across(f_id, ~mean(.x, na.rm = TRUE)))

# autumn average
foehn_y_sum_autumn <- foehn_d_sum |>
  mutate(winter =  as.numeric(format(date, "%m"))) |>
  filter(winter > 8 & winter < 12) |>
  mutate(year =  as.numeric(format(date, "%Y"))) |>
  group_by(year) |>
  summarize(across(f_id, ~mean(.x, na.rm = TRUE)))

foehn_y_sum$winter <- foehn_y_sum_winter$f_id
foehn_y_sum$summer <- foehn_y_sum_summer$f_id
foehn_y_sum$spring <- foehn_y_sum_spring$f_id
foehn_y_sum$autumn <- foehn_y_sum_autumn$f_id

```


### Aim of this document

In this report I investigate the foehn wind data on a basic level to get an understanding of the structure etc. In this analysis, all 8 stations are analyzed together.

<br>

#### Seasonality and distribution of foehn wind:
- we have a spring peak
- very little foehn hours in summer
- average in autumn and winter
- low day intensities are dominating
- mean daily foehn wind is 79.6 (18.51 when we include 0 foehn wind days)
- median daily foeh wind is 59  (0 when we include 0 foehn wind days)

```{r trends and seasonality, echo=FALSE, fig.height=3.5, fig.width=10, fig.align='center'}
# monthly plot
p1 = ggplot(foehn_m_mean, aes(x = month, y = f_id)) +
  geom_line(color = "red3", lwd=1.2) +
  theme_classic() +
  scale_color_viridis(discrete = TRUE) +
  labs(title = "average monthly foehn wind",
       x = "month",
       y = "foehn wind") +
  scale_x_continuous(breaks = unique(foehn_m_mean$month))


p2 = ggplot(foehn_d_sum[which(foehn_d_sum$f_id != 0),], aes(x = f_id))+
  geom_histogram(color = "black", fill = "red3", binwidth = 15, alpha = 0.7 )+
  theme_classic()+
  labs(title= "average daily foehn wind (without 0-foehn days)", x = "daily foehn wind")

grid.arrange(p1,p2,ncol=2)
```


```{r average daily foehn intensity without no foehn events, echo=FALSE, fig.align='center', fig.width=9, fig.height=5}
foehn_d_mean <- data[data$f_id != 0,] |>
  group_by(daymonth) |>
  summarize(across(f_id, ~ mean(.x, na.rm = TRUE))) |>
  mutate(date = paste0("2000.", daymonth),
         date = as.Date(date, format = "%Y.%m.%d"))

foehn_d_median <- data[data$f_id != 0,]  |>
  group_by(daymonth) |>
  summarize(across(f_id, ~ median(.x, na.rm = TRUE)))
foehn_d_median$date <- foehn_d_mean$date

plot(foehn_d_mean$date, foehn_d_mean$f_id, 
     xlab = "Day of the year", ylab = "mean daily foehn wind", col = "black", pch =1)
moving_mean <-c(NA,NA, NA,  rollmean(foehn_d_mean$f_id, k = 7), NA, NA,NA)
lines(foehn_d_mean$date, moving_mean, col = "brown1", lwd = 2.5)


moving_med <-c(NA,NA, NA,  rollmean(foehn_d_median$f_id, k = 7), NA, NA,NA)
lines(foehn_d_mean$date, moving_med, col = "brown4", lwd = 2.5)
legend("top", legend = c("7-day MA mean", "7-day MA median"), 
       col = c("brown1", "brown4"), lwd = 2, bty = "n")
```

<br>

#### Trends
- variability change in 2008 (due to different MedStat regions and start of observations for some stations)
- if we correct for the mean differences for the sections before and after 2008 we have a decreasing tendency for hospitalizations
- spring is stationary
- all other seasons slightly decreasing

```{r trends, echo=FALSE, fig.height=3.5}

# trend plots
foehn_y_long = foehn_y_sum[,c(1:6)] %>%
  pivot_longer(cols = c(f_id: autumn),
               names_to = "variable", 
               values_to = "value")

ggplot(foehn_y_long, aes(x = year, y = value, color = variable)) +
  geom_line(lwd = 1) +
  theme_classic() +
  scale_color_viridis(discrete = TRUE, labels = c("all", "Dec-Feb", "Mar-May", "Jun-Aug", "Sep-Nov")) +
  labs(title = "Yearly foehn score",
       x = "year",
       y = "") +
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "right")



mean1 = mean(foehn_y_sum$f_id[1:10])
mean2 = mean(foehn_y_sum$f_id[11:22])
diff = mean2-mean1
foehn_y_sum$f_id_corr = foehn_y_sum$f_id
foehn_y_sum$f_id_corr[11:22] = foehn_y_sum$f_id_corr[11:22] - diff

foehn_y_long = foehn_y_sum[,c(1,7)] %>%
  pivot_longer(cols = c(f_id_corr),
               names_to = "variable", 
               values_to = "value")
  
ggplot(foehn_y_long, aes(x = year, y = value, color = variable)) +
  geom_line(lwd = 1) +
  theme_classic() +
  scale_color_viridis(discrete = TRUE, labels = c("all", "Dec-Feb", "Mar-May", "Jun-Aug", "Sep-Nov")) +
  labs(title = "Yearly foehn score corrected",
       x = "year",
       y = "") +
  theme(legend.title = element_blank(),
        legend.text = element_text(size = 8),
        legend.position = "right")




```

<br>







