---
title: "Stations and their contribution to hospitalizations"
author: "Tino Schneidewind"
date: "`r Sys.Date()`"
output: html_document
---


## Aim of this document

Provide an overview over the available hospitalization counts and their distribution stratified by the meteorological stations that the Medstats were aggregated to. This investigates the reliability of hospitalizations and therefore whether there is a sufficient signal from the hospitalizations because these stations were originally chosen based on the foehn winds availability. The buffer radius by which the Medstats were aggregated is 8km. 



```{r load data, echo=FALSE, message=FALSE, warning=FALSE}
# clean environment
rm(list = ls())

# libraries
library(tidyverse); library(kableExtra)

# load data
data = read.csv("../../data/Medstat_hospitalizations_aggregated/hosp_buffer_8000.csv")
```



```{r loop mean, echo=FALSE}
tot_all = sum(data$all)

data_station <- data |>
  group_by(station)|>
  summarize(all_mean = mean(all, na.rm = TRUE),
            all_median = median(all, na.rm = TRUE),
            all_sd = sd(all, na.rm = TRUE),
            all_mean_sd_ratio = mean(all, na.rm = TRUE) / sd(all, na.rm = TRUE),
            all_percentage_contribution = (sum(all, na.rm = TRUE) / tot_all) * 100 )


knitr::kable(data_station, 
             col.names = c("Station","Mean","Median","SD", "MEAN/SD", "Contribution in %"),
             digits = 2,
             caption = "Table 1: Metrics for all cause hospitalizations per meteorological station.", 
             align = rep("l",4)) |>
  kable_styling(full_width = FALSE, position = "center")
```

Station experience different average daily hospitalizations e.g. less than one in Poschiavo and Davos, while Magadino has 5 and Lugano 12. The mean to sd ratio is relatively comparable between all stations. Apart from the previously discussed outliers, the contribution to total all cause hospitalization is comparable, with the exception of Lugano, which contributes 36%.



```{r dist, echo=FALSE, fig.height=9, fig.width=7, fig.align='center'}

ggplot(data = data) +
  geom_histogram(aes(x = all), bins = 60, fill = "steelblue", color = "black") +
  facet_wrap(~station, scales = "free_y", ncol = 2) +
  coord_cartesian(xlim = c(0, 25)) +
  theme_minimal()
```
<center>
Figure 1: Histograms of all cause daily hospitalizations stratified by station.
</center>

<br>
