---
title: "descriptive statisitcs: figures and tables"
author: "Tino Schneidewind"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data, echo=FALSE, warning=FALSE, message=FALSE}
## Preamble ####
rm(list=ls())

# packages
library(dplyr); library(tidyr); library(plotly);library(zoo); library(viridis); library(knitr);library(kableExtra);library(webshot2); library(magick)

data = read.csv("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/data/MedStat_aggregated/centroid_aggregated/hosp_buffer_8000.csv") # data

# BINARY FOEHN DEFINITION
data$f_id_binary <- ifelse(data$f_id >= 72, 1, 0)

data$date = as.Date(data$date); data$station <- as.factor(data$station)

# index to include only stratum that have hosp counts
data$stratum_dow = as.factor(data$stratum_dow); data$stratum = as.factor(data$stratum)
ind_dow = tapply(data$all, data$stratum_dow, sum); ind = tapply(data$all, data$stratum, sum)

data <- data %>%
  mutate(y64 = a014y + a1564y) %>%
  mutate(o64 = a6574y + a7584y + a85plusy)

# create uninform timeseries of means
data_time_serie = data %>%
  group_by(date) %>%
  summarise(
    mean_f_id = mean(f_id, na.rm = TRUE),
    mean_temp = mean(temp, na.rm = TRUE),
    mean_all = mean(all, na.rm = TRUE))

# moving averages for time series
data_time_serie$MA_f_id = rollmean(data_time_serie$mean_f_id, k = 30, fill = NA, align = "center")
data_time_serie$MA_all = rollmean(data_time_serie$mean_all, k = 30, fill = NA, align = "center")
data_time_serie$MA_temp = rollmean(data_time_serie$mean_temp, k = 30, fill = NA, align = "center")

# start dates of station records
start_dates = data.frame(station = as.vector(unique(data$station)),
                         date = rep("no",8))

for (i  in 1:8) {
  subs = data[data$station == as.character(unique(data$station)[i]),]
  start_dates$date[i] <- as.character(min(subs$date))
  
}
start_dates$date <- as.Date(start_dates$date)

# daily means
data_daily_mean = data %>%
  mutate(daymonth = format(date, "%m-%d")) %>%
  group_by(daymonth) %>%
  summarise(
    across(c(mal, fem, y64, o64, cvd, resp), mean),
    mean_f_id = mean(f_id),
    mean_temp = mean(temp),
    mean_all = mean(all),
    p50_f_id = quantile(f_id, 0.5),
    p95_f_id = quantile(f_id, 0.95),
    p75_f_id = quantile(f_id, 0.8),
    p90_f_id = quantile(f_id, 0.9),
    p25_temp = quantile(temp, 0.25),
    p50_temp = quantile(temp, 0.5),
    p75_temp = quantile(temp, 0.75),
    p25_all = quantile(all, 0.25),
    p50_all = quantile(all, 0.5),
    p75_all = quantile(all, 0.75),
  ) %>%
  mutate(daymonth = as.Date(paste0("2000-", daymonth)))

# moving averages of daily data
data_daily_mean$MA50 = rollmean(data_daily_mean$p50_f_id, k = 30, fill = NA, align = "center")
data_daily_mean$MA95 = rollmean(data_daily_mean$p95_f_id, k = 30, fill = NA, align = "center")
data_daily_mean$MA75 = rollmean(data_daily_mean$p75_f_id, k = 30, fill = NA, align = "center")
data_daily_mean$MA90 = rollmean(data_daily_mean$p90_f_id, k = 30, fill = NA, align = "center")

data_daily_mean$MA25_temp = rollmean(data_daily_mean$p25_temp, k = 30, fill = NA, align = "center")
data_daily_mean$MA50_temp = rollmean(data_daily_mean$p50_temp, k = 30, fill = NA, align = "center")
data_daily_mean$MA75_temp = rollmean(data_daily_mean$p75_temp, k = 30, fill = NA, align = "center")

data_daily_mean$MA25_all = rollmean(data_daily_mean$p25_all, k = 30, fill = NA, align = "center")
data_daily_mean$MA50_all = rollmean(data_daily_mean$p50_all, k = 30, fill = NA, align = "center")
data_daily_mean$MA75_all = rollmean(data_daily_mean$p75_all, k = 30, fill = NA, align = "center")

# ticks annual cycle x axis
monthly_ticks <- data_daily_mean$daymonth[!duplicated(format(data_daily_mean$daymonth, "%Y-%m"))]

# colors
colors <- viridis(3, option = "viridis")

# for Medstat table
station_MedStat = read.csv("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/data/MetStatRegions/centroids/per_station/MDSTID_MetRegions_8000_per_station.csv", header = TRUE)

station_MedStat_full <- rbind(c("ALT","CHU","DAV","LUG","MAG","MVE","ROB","VIS"), #Abbreviation
                              c(46.887042,46.870622,46.812956,46.003833,46.160019,46.298772,46.347194,46.302875),  #x coord
                              c(8.621806,9.530814,9.843492,8.960117,8.933608,7.460761,10.062964,7.842903),  #y coord
                              station_MedStat)

rownames(station_MedStat_full) <- c("Abbreviation", "X Coordinate", "Y Coordinate", "1 MedStat regions", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
# end of Medstat Table

# for calculating annual station averages
data$station_year = paste0(data$station, data$year)

# for station tables
stations = as.vector(unique(data$station)) 


```

<br>

This Markdown document presents some descriptive statistics about the key variables: foehn, temperature and hospitalizations. The focus lies on the distribution, the annual cycle and for hospitalizations the percentages of the subgroups.The shown data is generated by using the 8km buffer around the measurement stations.

<br>

#### Results

*Table 1: Descriptive statistics for the whole data set and for every station. Temperature statistics are based on annual means in °C. Foehn statistics are based on annual sums and the percentage indicates the ratio of the sum of foehn per station  compared to the sum of foehn from all stations. The binary foehn percentage shows the percentage of days with foehn with a binary threshold of 72 which corresponds to 6 h of full foehn. Hospitalization characteristics are based on annual sums. The percentage of hospitalizations indicates the contribution of hospitalizations per station to the whole data set. NA corresponds to the missing days. Standard deviations are given in brackets.*

```{r gen_table_station, echo=FALSE}
gen_stats = data.frame(period = rep(NA,9),
                       years = rep(NA,9),
                       temp = rep(NA, 9),
                       foehn = rep(NA,9),
                       per_tot_foehn = rep(NA,9),
                       foehn_bin_perc = rep(NA,9),
                       hosp = rep(NA,9),
                       per_tot_hosp = rep(NA,9),
                       missing_data = rep(NA,9),
                       # years = rep(NA,9),
                       # years = rep(NA,9),
                       # years = rep(NA,9),
                       row.names = c("all stations",stations)
                       # row.names = c("ALL","ALT", "CHU", "DAV", "LUG", "MAG", "MVE", "ROB", "VIS")
                       )

tot_foehn = sum(data$f_id)
tot_hosp = sum(data$all)



# yearly mean data weather
subs_y_all_weather = data %>%
  group_by(station_year) %>%
  summarize(mittel_temp = mean(temp, na.rm = TRUE),
            summe_foehn = sum(f_id, na.rm = TRUE)) 

# yearly mean data hosp
subs_y_all_hosp = data %>%
  group_by(station_year) %>%
  summarize(summe = sum(all, na.rm = TRUE))

# gen_stats$period[1] = paste0(min(data$date), " - ", max(data$date))
gen_stats$period[1] = as.character(min(data$date))
gen_stats$years[1] = length(unique(data$station_year))
gen_stats$temp[1] = paste0( round(mean(subs_y_all_weather$mittel_temp),digits=1), " [", round(sd(subs_y_all_weather$mittel_temp), digits= 1), "]")
gen_stats$foehn[1] = paste0( round(mean(subs_y_all_weather$summe_foehn),digits=0), " [", round(sd(subs_y_all_weather$summe_foehn), digits= 0), "]")
gen_stats$per_tot_foehn[1] = "100 %"
gen_stats$foehn_bin_perc[1] = paste0(round(length(which(data$f_id_binary>0))/nrow(data), digits = 3) *100, " %")
gen_stats$hosp[1] = paste0( round(mean(subs_y_all_hosp$summe),digits=0), " [", round(sd(subs_y_all_hosp$summe), digits= 0), "]")
gen_stats$per_tot_hosp[1] = "100 %"


# missing days
counter_mis = 0
counter_tot = 0

for(i in as.vector(unique(data$station_year))){
  subs = data[data$station_year == i,]
  
  full_dates = seq(min(subs$date), max(subs$date), by = "day")
  counter_tot = counter_tot + length(full_dates)
  counter_mis = counter_mis + ( length(full_dates) - nrow(subs) )
  
}

gen_stats$missing_data[1] = paste0( round(counter_mis/counter_tot, digits = 3) *100, " %")

for (i in 1:length(stations)) {
  sta = stations[i]
  
  subs = data[data$station == sta,]
  
  # yearly mean data weather
  subs_y_all_weather = subs %>%
    group_by(station_year) %>%
    summarize(mittel_temp = mean(temp, na.rm = TRUE),
             summe_foehn = sum(f_id, na.rm = TRUE)) 

  # yearly mean data hosp
  subs_y_all_hosp = subs %>%
    group_by(station_year) %>%
    summarize(summe = sum(all, na.rm = TRUE))

  # gen_stats$period[i+1] = paste0(min(subs$date), " - ", max(subs$date))
  gen_stats$period[i+1] = as.character(min(subs$date))
  gen_stats$years[i+1] = length(unique(subs$station_year))
  gen_stats$temp[i+1] = paste0( round(mean(subs_y_all_weather$mittel_temp),digits=1), " [", round(sd(subs_y_all_weather$mittel_temp), digits= 1), "]")
  gen_stats$foehn[i+1] = paste0( round(mean(subs_y_all_weather$summe_foehn),digits=0), " [", round(sd(subs_y_all_weather$summe_foehn), digits= 0), "]")
  gen_stats$per_tot_foehn[i+1] =  paste0( round( sum(subs$f_id) / tot_foehn, digits=3) *100, " %")
  gen_stats$foehn_bin_perc[i+1] = paste0(round(length(which(subs$f_id_binary>0))/nrow(subs), digits = 3) *100, " %")
  gen_stats$hosp[i+1] = paste0( round(mean(subs_y_all_hosp$summe),digits=0), " [", round(sd(subs_y_all_hosp$summe), digits= 0), "]")
  gen_stats$per_tot_hosp[i+1] = paste0(round(sum(subs$all) / tot_hosp, digits=3) *100, " %")


  full_dates = seq(min(subs$date), max(subs$date), by = "day")
  gen_stats$missing_data[i+1] = paste0( round(  length(full_dates) / nrow(subs) *100 - 100 , digits = 1), " %")

  }


kable(gen_stats,
      col.names = c("start date", "years", "temp", "foehn" ,"%-foehn", "bin-foehn-%", "hosp", "%-hosp", "NA"      ),
      align = c("c", rep("r", ncol(gen_stats) -2), "c"))  #%>%
#    kable_styling() %>%
# save_kable("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/main_table.html")
# webshot("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/main_table.html",
#         "C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/main_table.png",
#         vwidth = 1000,   # increase width
#         vheight = 400,  # increase height
#         zoom = 3)        # increase zoom level for higher resolution
```

<br>

#### Appendix


```{r plot_foehn, echo=FALSE, fig.width= 6, fig.height=6, fig.align = 'center'}

png("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/distribution_plot.png", width = 1800, height = 2000, res = 300)

par(mfrow=c(3,2), 
    mar = c(4,4,.5,1),
    mgp = c(2, .5, 0))


hist(data$f_id[data$f_id!=0], 
     breaks = 40,
     col = colors[2], 
     xlim = c(0,300),
     xlab = "foehn", 
     main = "", 
     # cex.axis = 0.6,
     ylab = "frequency")
text(285, 2250, labels = "(a)", pos = 2)

plot(data_daily_mean$daymonth, 
     data_daily_mean$mean_f_id, 
     xaxt = "n", col = colors[2],
     type = "p",
     xlab = "month", 
     ylab = "foehn", 
     main = "",
     ylim = c(0,140), 
     pch = 16, 
     cex = .6,
     # cex.axis = 0.6, 
     # lwd = 2,
     bty = "n")

axis(1, 
     at = monthly_ticks + 15, 
     labels = substr(format(monthly_ticks, "%b"),1,1)#,cex.axis = 0.6
     )

lines(data_daily_mean$daymonth, data_daily_mean$MA50, col = 1, lwd = 2, lty = 1)
lines(data_daily_mean$daymonth, data_daily_mean$MA75, col = 1, lwd = 2, lty = 2)
lines(data_daily_mean$daymonth, data_daily_mean$MA90, col = 1, lwd = 2, lty = 3)
# lines(data_daily_mean$daymonth, data_daily_mean$MA95, col = 5, lwd = 2, lty = 1)

legend(monthly_ticks[6]+15, 120,
       legend = c("p50", "p75","p90"), bty = "n",
       col = c(1,1,1) ,lwd = 1, lty = c(1,2,3), cex = 0.8, 
       ncol = 2)

text(monthly_ticks[12], 135, 
     labels = "(b)", pos = 2)

hist(data$temp, 
     breaks = 30,
     col = colors[1], 
     # xlim = c(0,300),
     xlab = "temperature [\u00B0C]", 
     main = "", 
     # cex.axis = 0.6,
     ylab = "frequency")
text(28, 3950, labels = "(c)", pos = 2)

plot(data_daily_mean$daymonth, 
     data_daily_mean$mean_temp, 
     xaxt = "n", col = colors[1],
     type = "p",
     xlab = "month", 
     ylab = "temperature [\u00B0C]", 
     main = "",
     ylim = c(-5,25),
     pch = 16, 
     cex = .6,
     # cex.axis = 0.6, 
     bty = "n")

axis(1, 
     at = monthly_ticks + 15, 
     labels = substr(format(monthly_ticks, "%b"),1,1)#, cex.axis = 0.6
     )

lines(data_daily_mean$daymonth, data_daily_mean$MA25_temp, col = 1, lwd = 2, lty = 3)
lines(data_daily_mean$daymonth, data_daily_mean$MA50_temp, col = 1, lwd = 2, lty = 1)
lines(data_daily_mean$daymonth, data_daily_mean$MA75_temp, col = 1, lwd = 2, lty = 2)

legend(monthly_ticks[2]+15, -2,  
       legend = c("p25", "p50","p75"), bty = "n",
       col = c(1,1,1),lwd = 1, lty = c(3,1,2), cex = 0.8, 
       ncol = 3)

text(monthly_ticks[12], 23, labels = "(d)", pos = 2)


hist(data$all, 
     breaks = 50,
     col = colors[3], 
     xlim = c(0,20),
     xlab = "hospitalizations", 
     main = "", 
     # cex.axis = 0.6,
     ylab = "frequency")
text(19, 21000, labels = "(e)", pos = 2)

plot(data_daily_mean$daymonth, 
     data_daily_mean$mean_all, 
     xaxt = "n", col = colors[3],
     type = "p",
     xlab = "month", 
     ylab = "hospitalizations", 
     main = "",
     ylim = c(0,6),
     pch = 16, 
     cex = .6,
     # cex.axis = 0.6, 
     bty = "n")

axis(1, 
     at = monthly_ticks + 15, 
     labels = substr(format(monthly_ticks, "%b"),1,1)#, cex.axis = 0.6
     )

lines(data_daily_mean$daymonth, data_daily_mean$MA25_all, col = 1, lwd = 2, lty = 3)
lines(data_daily_mean$daymonth, data_daily_mean$MA50_all, col = 1, lwd = 2, lty = 1)
lines(data_daily_mean$daymonth, data_daily_mean$MA75_all, col = 1, lwd = 2, lty = 2)

legend(monthly_ticks[3]+5, 5.75,  
       legend = c("p25", "p50","p75"), bty = "n",
       col = c(1,1,1),lwd = 1, lty = c(3,1,2), cex = 0.8, 
       ncol = 3)

text(monthly_ticks[12], 6, labels = "(f)", pos = 2)

dev.off()
```

*Figure 1: (a) daily foehn score distribution excluding 0-foehn days, (b) daily mean foehn score with 30-day moving averages of the 50th, 75th, 90th percentile, (c) daily mean temperature distribution, (d) daily averages of daily mean temperature with 30-day moving averages of the 25th, 50th, 75th percentile, (e) daily all-cause hospitalization distribution, (f) daily mean all-cause hospitalizations with 30-day moving averages of the 25th, 50th, 75th percentile.*

<br>


```{r timeseriesplot, echo=FALSE, fig.width= 5, fig.height=4.5, fig.align = 'center'}
png("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/timeseries_plot.png", width = 1600, height = 1400, res = 300)

# margins
par(mfrow = c(3, 1), 
    mar = c(0, 4, 0, 0.5),  # Top two plots have zero bottom margin
    oma = c(4, 0, 0, 0),
    mgp = c(2, .5, 0))      # Outer bottom margin for x-axis label

# foehn plot
plot(data_time_serie$date, data_time_serie$MA_f_id,
     type = "n",
     xlab = "",
     ylab = "foehn",
     xaxt = "n"
     )

for(i in start_dates$date){  abline(v = i, col = 1, lty = 2)}

abline(v = as.Date("2008-01-01"), col = "brown1", lty = 1)

lines(data_time_serie$date, data_time_serie$MA_f_id,     
      lty = 1,
     lwd = 2,
     col = colors[2],)

text(data_time_serie$date[nrow(data_time_serie)], 78, labels = "(a)", pos = 2)

# temp plot
plot(data_time_serie$date, data_time_serie$MA_temp,
     type = "n",
     xlab = "",
     ylab = "temperature [\u00B0C]",
     xaxt = "n"
     )

for(i in start_dates$date){  abline(v = i, col = 1, lty = 2)}

abline(v = as.Date("2008-01-01"), col = "brown1", lty = 1)

lines(data_time_serie$date, data_time_serie$MA_temp,
      lty = 1,
     lwd = 2,
     col = colors[1],)

text(data_time_serie$date[nrow(data_time_serie)], -2, labels = "(b)", pos = 2)

# hosp plot
plot(data_time_serie$date, data_time_serie$MA_all,
     type = "n",
     xlab = "",
     ylab = "hospitalizations",
     )

mtext("time", side = 1, line = 2, outer = TRUE, cex = 0.8)

for(i in start_dates$date){  abline(v = i, col = 1, lty = 2)}

abline(v = as.Date("2008-01-01"), col = "brown1", lty = 1)

lines(data_time_serie$date, data_time_serie$MA_all,      
     lty = 1,
     lwd = 2,
     col = colors[3])

text(data_time_serie$date[nrow(data_time_serie)], 1.8, labels = "(c)", pos = 2)

text(start_dates$date[1]+85, 4.8, labels = "(1)", pos = 2, cex = .8)
text(start_dates$date[4]+125, 4.8, labels = "(3)", pos = 2, cex = .8)
text(start_dates$date[5]+85, 4.8, labels = "(2)", pos = 2, cex = .8)
text(start_dates$date[7]-70, 4.8, labels = "(5)", pos = 4, cex = .8)
text(start_dates$date[8]+85, 4.8, labels = "(4)", pos = 2, cex = .8)
text(as.Date("2008-01-01")+70, 5.2, labels = "(x)", pos = 2, col = "brown1", cex = .8)

dev.off()
  
```

*Figure 2: (a) 30-day moving average time series of mean daily foehn, (b) 30-day moving average time series of average daily mean temperature, (c) 30-day moving average of mean daily all-cause hospitalization. (1) start of data from Altdorf, Chur, Davos, Montana, (2) start of data from Magadino, (3) start of data from Lugano, (4) start of data from Visp, (5) start of data from Poschiavo. (x) redefinition of the MedStat regions by the Swiss Federal Office for Statistics.*

<br>

*Table 2: Contribution in percent of different subgroups to all-cause hospitalizations counts across all stations and per station.*



```{r subgroubs_table_stations, echo=FALSE}

subgroups = data.frame(mal = rep(NA, 9), 
                       row.names = c("all stations", stations)) 


subgroups$mal[1] = paste( round( sum(data$mal) / sum(data$all), digits = 4) *100, " %")
subgroups$fem[1] = paste( round( sum(data$fem) / sum(data$all), digits = 4) *100, " %")

subgroups$y64[1] = paste( round( sum(data$y64) / sum(data$all), digits = 4) *100, " %")
subgroups$o64[1] = paste( round( sum(data$o64) / sum(data$all), digits = 4) *100, " %")

subgroups$cvd[1] = paste( round( sum(data$cvd) / sum(data$all), digits = 4) *100, " %")
subgroups$resp[1] = paste( round( sum(data$resp) / sum(data$all), digits = 4) *100, " %")
subgroups$inf[1] = paste( round( sum(data$inf) / sum(data$all), digits = 4) *100, " %")
subgroups$uri[1] = paste( round( sum(data$uri) / sum(data$all), digits = 4) *100, " %")
subgroups$ment[1] = paste( round( sum(data$ment) / sum(data$all), digits = 4) *100, " %")


for (i in 1:length(unique(data$station))) {
  subs = data[data$station == unique(data$station)[i],]
  
  subgroups$mal[i+1] = paste( round( sum(subs$mal) / sum(subs$all), digits = 4) *100, " %")
  subgroups$fem[i+1] = paste( round( sum(subs$fem) / sum(subs$all), digits = 4) *100, " %")

  subgroups$y64[i+1] = paste( round( sum(subs$y64) / sum(subs$all), digits = 4) *100, " %")
  subgroups$o64[i+1] = paste( round( sum(subs$o64) / sum(subs$all), digits = 4) *100, " %")

  subgroups$cvd[i+1] = paste( round( sum(subs$cvd) / sum(subs$all), digits = 4) *100, " %")
  subgroups$resp[i+1] = paste( round( sum(subs$resp) / sum(subs$all), digits = 4) *100, " %")
  subgroups$inf[i+1] = paste( round( sum(subs$inf) / sum(subs$all), digits = 4) *100, " %")
  subgroups$uri[i+1] = paste( round( sum(subs$uri) / sum(subs$all), digits = 4) *100, " %")
  subgroups$ment[i+1] = paste( round( sum(subs$ment) / sum(subs$all), digits = 4) *100, " %")
  
  
}                       


kable(subgroups, 
      digits = 2, 
      col.names = c("male", "female", "age <= 64", "age > 64", "cvd", "resp", "inf", "uri", "ment"),
      align = rep("r", ncol(subgroups))) #%>%
#    kable_styling() %>%
# save_kable("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/subgroups.html")
# webshot("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/subgroups.html",
#         "C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/subgroups_table.png",
#         vwidth = 1000,   # increase width
#         vheight = 400,  # increase height
#         zoom = 3)        # increase zoom level for higher resolution

```



<br>

*Table 3: Abbreviation and location of meteorological measurement stations and their assigned MedStat regions within the 8 km buffer.*


```{r station_stats, echo=FALSE}


kable(station_MedStat_full,
      align = rep("r", ncol(station_MedStat_full))) # %>%
#    kable_styling() %>%
# save_kable("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/medstat_table.html")
# webshot("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/medstat_table.html",
#         "C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/medstat_table.png",
#         vwidth = 1000,   # increase width
#         vheight = 400,  # increase height
#         zoom = 3)        # increase zoom level for higher resolution
```

<br>
