---
title: "the direct effect of foehn on hospitalizations"
author: "Tino Schneidewind"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r data, echo=FALSE, warning=FALSE, message=FALSE}
### PACKAGES ####
library(dlnm);library(splines);library(ggplot2);library(viridis);library(gnm); library(dplyr);library(knitr);library(lmtest)

rm(list=ls())



### DATA ####

# buffer size for data file read
buffer = 8000

data = read.csv(paste0("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/data/MedStat_aggregated/centroid_aggregated/hosp_buffer_", buffer, ".csv"))

data$date = as.Date(data$date)
data$station <- as.factor(data$station)

# index to include only stratum that have hosp counts
data$stratum_dow = as.factor(data$stratum_dow); data$stratum = as.factor(data$stratum)
ind_dow = tapply(data$all, data$stratum_dow, sum); ind = tapply(data$all, data$stratum, sum)


# create larger age groups
data <- data %>%
  mutate(y64 = a014y + a1564y) %>%
  mutate(o64 = a6574y + a7584y + a85plusy)

# define the maximum lag distance we account for
maxlago <- 3

# mmt function
source("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/master_thesis/functions/findmin.R")

# crossbasis temp
cb.temp <- crossbasis(data$temp,
                      lag=21,
                      argvar=list(fun="ns", knots = quantile(data$temp, c(.5,.9), na.rm=TRUE)),
                      arglag=list(fun="ns", knots = logknots(21,3)),
                      group = data$station)

# binary foehn
foehn_bin <- ifelse(data$f_id > 72, 0, 1)
foehn_bin_rev <- ifelse(foehn_bin == 1, 0, 1)

# modifier functions
modif <- cb.temp * foehn_bin
modif_rev <- cb.temp * foehn_bin_rev

# groups
groups_id = colnames(data)[c(3,9,10,24,25, 13, 14, 11, 15, 12)]

letters = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j")


# colors
colors <- viridis(3, option = "viridis")

foehn_col = colors[2]
temp_col = colors[3]


# seasonal data
## determine months for both warm and cold season 
monthly_means = data %>% 
  group_by(month) %>%
  summarize(mean_temp = mean(temp, na.rm=TRUE))

index_s = which(monthly_means$mean_temp > median(monthly_means$mean_temp))
index_w = which(monthly_means$mean_temp < median(monthly_means$mean_temp))

data_summer = data[data$month %in% index_s,]
data_winter = data[data$month %in% index_w, ]

data_summer$station_year = paste0(data_summer$station, data_summer$year)
data_winter$station_year = paste(data_winter$station, ifelse(data_winter$month %in% 1:10, data_winter$year, data_winter$year+1))

#####
```

<br>

This Markdown document presents and summarizes the findings of an analysis concerning the following research question.

<br>

#### *Results for RQ2+RQ3: is the temperature-hospitalizations association significantly modified by foehn in Switzerland? are certain subgroups more affected than others?*

<br>


```{r results_all, echo=FALSE, fig.width=3.3, fig.height=3.3, fig.align='center'}
png("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/rq3_all_plot.png", width = 1000, height =1000, res = 300)

par(mfrow=c(1,1), 
    mar = c(3,3,.5,.5),
    mgp = c(1.8, .5, 0))

foehn_col = "green4"
# temp_col = "purple"
temp_col = "gold2"

for (i in 1:length(groups_id[1])) {

  colvar  =  groups_id[i]

  formula1  <- as.formula(paste0(colvar, "~ cb.temp + modif"))
  formula2 <- as.formula(paste0(colvar, "~ cb.temp + modif_rev"))

  # model with and without foehn
  mod_modif     <-gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)
  mod_modif_rev <- gnm(formula2, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # prediction with and without foehn
  pred_modif   <- crosspred(cb.temp, mod_modif, cen = 20)
  pred_modif2  <- crosspred(cb.temp, mod_modif_rev, cen = 20)

  
  # get min value of both predictions and use for centering
  min1 <- findmin(cb.temp,pred_modif,from=quantile(data$temp, .1),to=quantile(data$temp, .9))
  # min2 <- findmin(cb.temp,pred_modif2,from=quantile(data$temp, .1),to=quantile(data$temp, .9))

  # predict with new min values
  pred_modif_new     <- crosspred(cb.temp, mod_modif, cen = min1, by = .1)
  pred_modif_rev_new <- crosspred(cb.temp, mod_modif_rev, cen = min1, by = .1)
  
  plot(pred_modif_new,              ## cumulative exposure
     "overall",
     ylab = "cumulative RR",
     xlab = "temperature [\u00B0C]",
     col = foehn_col,
     ci.arg = list(col = alpha(colour = foehn_col, .15)),
     lwd = 2,
     main ="",
     ylim = c(0.7,2.5),
     cex.axis = 0.7,
     cex.lab = 0.7)

  
  lines(pred_modif_rev_new,           ## cumulative exposure
     "overall",
     col = temp_col,
     ci = "area",
     ci.arg = list(col = alpha(colour = temp_col, .15)),
     lwd = 2)
  
  abline(v =quantile(data$temp, .99), col = "black", lty = 2)
  
    legend("top", ncol = 1, legend = c("temp with interaction", "temp without interaction"), col = c("green4", "gold2"),
         bty = "n", lwd=c(2,2), cex = 0.7)
  
  text(25, 2.5 , labels = "(a)", pos = 4, cex = 0.7)
  }


dev.off()
```

```{r seasonal, echo=FALSE}

# png("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/rq3_all_season.png", width = 1700, height =1000, res = 300)


par(mfrow=c(1,2), 
    mar = c(3,3,.5,.5),
    mgp = c(1.8, .5, 0))
# summer plot
# binary foehn
foehn_bin_sum <- ifelse(data_summer$f_id > 72, 0, 1)
foehn_bin_rev_sum <- ifelse(foehn_bin_sum == 1, 0, 1)

# crossbasis temp
cb.temp_sum <- crossbasis(data_summer$temp,
                      lag=21,
                      argvar=list(fun="ns", knots = quantile(data_summer$temp, c(.5,.9), na.rm=TRUE)),
                      arglag=list(fun="ns", knots = logknots(21,3)),
                      group = data_summer$station_year)

# modifier functions
modif_sum <- cb.temp_sum * foehn_bin_sum
modif_rev_sum <- cb.temp_sum * foehn_bin_rev_sum



mod_modif_sum     <-gnm(all ~ cb.temp_sum + modif_sum, data = data_summer,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)
mod_modif_rev_sum <- gnm(all ~ cb.temp_sum + modif_rev_sum, data = data_summer,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

# prediction with and without foehn
pred_modif_sum   <- crosspred(cb.temp_sum, mod_modif_sum, cen = 20, cumul=FALSE)
pred_modif2_sum  <- crosspred(cb.temp_sum, mod_modif_rev_sum, cen = 20, cumul=FALSE)


# get min value of both predictions and use for centering
min1_sum <- findmin(cb.temp_sum, pred_modif_sum,from=quantile(data_summer$temp, .1),to=quantile(data_summer$temp, .9))


# predict with new min values
pred_modif_new_sum     <- crosspred(cb.temp_sum, mod_modif_sum, cen = min1_sum, cumul=FALSE, by = .2)
pred_modif_rev_new_sum <- crosspred(cb.temp_sum, mod_modif_rev_sum, cen = min1_sum, cumul=FALSE, by = .2)

plot(pred_modif_new_sum,              ## cumulative exposure
     "overall",
     ylab = "cumulative RR",
     xlab = "temperature [\u00B0C]",
     col = foehn_col,
     ci.arg = list(col = alpha(colour = foehn_col, .15)),
     lwd = 2,
     main ="",
     ylim = c(0.7,2.5),
     cex.axis = 0.7,
     cex.lab = 0.7)


  lines(pred_modif_rev_new_sum,           ## cumulative exposure
     "overall",
     col = temp_col,
     ci = "area",
     ci.arg = list(col = alpha(colour = temp_col, .15)),
     lwd = 2)
  
  legend("topleft", ncol = 1, legend = c("temp with interaction", "temp without interaction"), col = c("green4", "gold2"),
         bty = "n", lwd=c(2,2), cex = 0.7)

text(25, 2.5 , labels = "(a)", pos = 2, cex = 0.7)


# winter plot
# binary foehn
foehn_bin_win <- ifelse(data_winter$f_id > 72, 0, 1)
foehn_bin_rev_win <- ifelse(foehn_bin_win == 1, 0, 1)

# crossbasis temp
cb.temp_win <- crossbasis(data_winter$temp,
                      lag=21,
                      argvar=list(fun="ns", knots = quantile(data_winter$temp, c(.5,.9), na.rm=TRUE)),
                      arglag=list(fun="ns", knots = logknots(21,3)),
                      group = data_winter$station_year)

# modifier functions
modif_win <- cb.temp_win * foehn_bin_win
modif_rev_win <- cb.temp_win * foehn_bin_rev_win



mod_modif_win     <-gnm(all ~ cb.temp_win + modif_win, data = data_winter,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)
mod_modif_rev_win <- gnm(all ~ cb.temp_win + modif_rev_win, data = data_winter,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

# prediction with and without foehn
pred_modif_win   <- crosspred(cb.temp_win, mod_modif_win, cen = 20, cumul=FALSE)
pred_modif2_win  <- crosspred(cb.temp_win, mod_modif_rev_win, cen = 20, cumul=FALSE)


# get min value of both predictions and use for centering
min1_win <- findmin(cb.temp_win, pred_modif_win,from=quantile(data_winter$temp, .1),to=quantile(data_winter$temp, .9))


# predict with new min values
pred_modif_new_win     <- crosspred(cb.temp_win, mod_modif_win, cen = min1_win, cumul=FALSE, by = .2)
pred_modif_rev_new_win <- crosspred(cb.temp_win, mod_modif_rev_win, cen = min1_win, cumul=FALSE, by = .2)

plot(pred_modif_new_win,              ## cumulative exposure
     "overall",
     ylab = "cumulative RR",
     xlab = "temperature [\u00B0C]",
     col = foehn_col,
     ci.arg = list(col = alpha(colour = foehn_col, .15)),
     lwd = 2,
     main ="",
     ylim = c(0.7,2.5),
     cex.axis = 0.7,
     cex.lab = 0.7)


  lines(pred_modif_rev_new_win,           ## cumulative exposure
     "overall",
     col = temp_col,
     ci = "area",
     ci.arg = list(col = alpha(colour = temp_col, .15)),
     lwd = 2)

text(25, 2.5 , labels = "(b)", pos = 2, cex = 0.7)

# dev.off()


```


```{r results_subgroups, echo=FALSE, fig.width=9, fig.height=9, fig.align='center'}
# png("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/rq3_subgroups_plot.png", width = 1800, height =1800, res = 300)

# foehn_col = "green4"
# # temp_col = "purple"
# temp_col = "gold2"

par(mfrow=c(3,3), 
    mar = c(3,3,.5,.5),
    mgp = c(1.8, .5, 0))

# for (i in 1:length(groups_id[2:10])) {
for (i in 1:length(groups_id[2:10])) {

  colvar  =  groups_id[i+1]

  formula1  <- as.formula(paste0(colvar, "~ cb.temp + modif"))
  formula2 <- as.formula(paste0(colvar, "~ cb.temp + modif_rev"))

  # model with and without foehn
  mod_modif     <-gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)
  mod_modif_rev <- gnm(formula2, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # prediction with and without foehn
  pred_modif   <- crosspred(cb.temp, mod_modif, cen = 20, cumul=FALSE)
  pred_modif2  <- crosspred(cb.temp, mod_modif_rev, cen = 20, cumul=FALSE)

  
  # get min value of both predictions and use for centering
  min1 <- findmin(cb.temp,pred_modif,from=quantile(data$temp, .1),to=quantile(data$temp, .9))
  min2 <- findmin(cb.temp,pred_modif2,from=quantile(data$temp, .1),to=quantile(data$temp, .9))

  # predict with new min values
  pred_modif_new     <- crosspred(cb.temp, mod_modif, cen = min1, cumul=FALSE, by = .1)
  pred_modif_rev_new <- crosspred(cb.temp, mod_modif_rev, cen = min1, cumul=FALSE, by = .1)
  
  plot(pred_modif_new,              ## cumulative exposure
     "overall",
     ylab = "cumulative RR",
     xlab = "temperature [\u00B0C]",
     col = foehn_col,
     ci.arg = list(col = alpha(colour = foehn_col, .2)),
     lwd = 2,
     main ="",
     ylim = c(0.7,3)
     )

  
  lines(pred_modif_rev_new,           ## cumulative exposure
     "overall",
     col = temp_col,
     ci = "area",
     ci.arg = list(col = alpha(colour = temp_col, .2)),
     lwd = 2)
  
  text(25, 3, labels = paste0("(", letters[i], ")"), pos = 2)}
# max(pred_modif_new$allRRhigh)

# dev.off()
```



*Figure 1: Cumulative response ratios for (a) all-cause, (b) male, (c), female, (d) 64 years and younger, (e) older than 64 years, (f) cvd, (g) resp, (h) inf, (i) uri, (j) ment hospitalization with a binary foehn threshold value of 72 which corresponds to 6 h of full foehn. The green line shows the temperature hospitallization association when foehn was present, the yellow line when foehn was abscent.*



<br>


```{r results_subgroups_table, echo=FALSE}
table_estimates = data.frame(all = rep(NA, 2),
                             mal = rep(NA, 2),
                             fem = rep(NA, 2),
                             y64 = rep(NA, 2),
                             o64  = rep(NA, 2),
                             cvd = rep(NA, 2),
                             resp = rep(NA, 2),
                             inf = rep(NA, 2),
                             uri = rep(NA, 2),
                             ment = rep(NA, 2),
                             row.names = c("RR temp + foehn [CI]", "RR temp - foehn [CI]"))


print(quantile(data$temp, .99))

for (i in 1:length(groups_id[2:10])) {

  colvar  =  groups_id[i]
  
  

  formula1  <- as.formula(paste0(colvar, "~ cb.temp + modif"))
  formula2 <- as.formula(paste0(colvar, "~ cb.temp + modif_rev"))

  # model with and without foehn
  mod_modif     <-gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)
  mod_modif_rev <- gnm(formula2, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # prediction with and without foehn
  pred_modif   <- crosspred(cb.temp, mod_modif, cen = 20, cumul=FALSE)
  pred_modif2  <- crosspred(cb.temp, mod_modif_rev, cen = 20, cumul=FALSE)

  
  # get min value of both predictions and use for centering
  min1 <- findmin(cb.temp,pred_modif,from=quantile(data$temp, .1),to=quantile(data$temp, .9))
  min2 <- findmin(cb.temp,pred_modif2,from=quantile(data$temp, .1),to=quantile(data$temp, .9))

  # predict with new min values
  pred_modif_new     <- crosspred(cb.temp, mod_modif, cen = min1, cumul=FALSE, by = .1)
  pred_modif_rev_new <- crosspred(cb.temp, mod_modif_rev, cen = min1, cumul=FALSE, by = .1)
 
  # extract prediction for value 24.7 and save it
  table_estimates[1,i] = paste0( sprintf("%.3f",round(pred_modif_new$allRRfit["24.7"],digits=3)), 
                                         " [", 
                                         sprintf("%.3f",round(pred_modif_new$allRRlow["24.7"], digits=3)),
                                         "-", 
                                         sprintf("%.3f",round(pred_modif_new$allRRhigh["24.7"], digits = 3)), "]")
  
  table_estimates[2,i] = paste0( sprintf("%.3f",round(pred_modif_rev_new$allRRfit["24.7"],digits=3)), 
                                         " [", 
                                         sprintf("%.3f",round(pred_modif_rev_new$allRRlow["24.7"], digits=3)),
                                         "-", 
                                         sprintf("%.3f",round(pred_modif_rev_new$allRRhigh["24.7"], digits = 3)), "]")
  
  
   
}

table_estimates_transposed = t(table_estimates)

kable(table_estimates_transposed,
      # digits = 3,
      col.names = c("RR temp + foehn [CI]", "RR temp - foehn [CI]"),
      align = rep("r", ncol(table_estimates_transposed) )) #%>%
#    kable_styling(full_width = FALSE) %>%
# save_kable("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/rq3_table.html")
# webshot("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/rq3_table.html",
#         "C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/rq3_table.png",
#         vwidth = 450,   # increase width
#         vheight = 400,  # increase height
#         zoom = 3)        # increase zoom level for higher resolution
# 



```

```{r results_subgroups_lollipop_figure, echo=FALSE}
library(ggstance)
table_estimates = data.frame(categories = c("all","all", "mal", "mal","fem","fem","y64","y64","o64","o64",
                                            "cvd","cvd","resp","resp","inf","inf","uri","uri","ment","ment"),
                             model = rep(c("temp + foehn", "temp - foehn"), 10),
                             pred = rep(NA, 20),
                             CI_low = rep(NA, 20),
                             CI_high = rep(NA, 20))

foehn_col = "green4"
# temp_col = "purple"
temp_col = "gold2"

for (i in 1:nrow(table_estimates)) {

  colvar  =  table_estimates$categories[i]
  
  if( i %% 2 != 0) {
    formula1 = as.formula(paste0(colvar, "~ cb.temp + modif"))
  } else {
    formula1 = as.formula(paste0(colvar, "~ cb.temp + modif_rev")) 
  }
    
  
  
  # model with and without foehn
  mod_modif     <-gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # prediction with and without foehn
  pred_modif   <- crosspred(cb.temp, mod_modif, cen = 20, cumul=FALSE)

  # get min value of both predictions and use for centering
  min1 <- findmin(cb.temp,pred_modif,from=quantile(data$temp, .1),to=quantile(data$temp, .9))

  # predict with new min values
  pred_modif_new  <- crosspred(cb.temp, mod_modif, cen = min1, cumul=FALSE, by = .1)
  
  
  table_estimates$pred[i]    = pred_modif_new$allRRfit["24.7"]
  table_estimates$CI_low[i]  = pred_modif_new$allRRlow["24.7"]
  table_estimates$CI_high[i] = pred_modif_new$allRRhigh["24.7"]


}
table_estimates$categories[7:8] = "y64"
table_estimates$categories[9:10] = "o64"


table_estimates$categories <- factor(table_estimates$categories, 
                                     levels = rev(unique(table_estimates$categories)))

table_estimates$model <- factor(table_estimates$model, 
                                levels = c("temp + foehn", "temp - foehn"))



# png("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/rq3_subgroups_lolli.png", 
    # width = 1200, height =1400, res = 300)
png("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/rq3_subgroups_lolli_new.png",
    width = 1200, height =1400, res = 300)

ggplot(table_estimates, aes(y = categories, x = pred, color = model)) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  geom_pointrange(aes(xmin = CI_low, xmax = CI_high), 
                  position = position_dodgev(height = 0.4)) +
  labs(title = "",
       x = "cumulative RR", y = "") +
  theme_minimal() +
  scale_color_manual(
    values = c("temp + foehn" = foehn_col, "temp - foehn" = temp_col), 
    labels = c("temp + foehn" = "temp with interaction", "temp - foehn" = "temp without interaction"),  # Map the levels to custom labels
    name = "model:"  # Custom legend title
  ) +
  annotate("text", x = Inf, y = Inf, label = "(b)", 
           hjust = 1.2, vjust = 1.2, size = 3) +
  
  theme(
    legend.position = "none",
    # legend.position = c(1,1), # Position the legend at the top
    # legend.justification = c("right", "top"),
    # legend.title = element_text(size = 10),  # Font size for the legend title
    # legend.text = element_text(size = 8),    # Font size for the legend text
    axis.text.x = element_text(size = 8),    # Font size for x-axis labels
    axis.text.y = element_text(size = 8),    # Font size for y-axis labels
    axis.title.x = element_text(size = 10),  # Font size for x-axis title
    plot.margin = unit(c(0, 1, 0, 0), "cm"), # Adjust plot margins
    axis.line = element_line(color = "black"), # Add axis lines
    panel.grid.minor = element_blank(),       # Remove minor grid lines
    panel.grid.major.y = element_blank(),     # Remove horizontal grid lines
    panel.grid.major.x = element_blank(),  # Remove minor grid lines
    legend.margin = margin(.5, .5, .5, .5),
    legend.background = element_rect(fill = "white", color = "white"),
  ) +
  guides(color = guide_legend(ncol = 1))  # Set legend to one column

dev.off()




```

**Prediction of the interaction term**

for all cause

```{r results_all_interaction, echo=FALSE, fig.width=3.3, fig.height=3.3, fig.align='center'}
# png("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/rq2_all_interaction_plot.png", width = 1000, height =1000, res = 300)

par(mfrow=c(1,1), 
    mar = c(3,3,.5,.5),
    mgp = c(1.8, .5, 0))

foehn_col = "green4"
# temp_col = "purple"
temp_col = "gold2"


for (i in 1:length(groups_id[3])) {

  colvar  =  groups_id[i+9]

  formula1  <- as.formula(paste0(colvar, "~ cb.temp + modif_rev"))
  formula2 <- as.formula(paste0(colvar, "~ cb.temp")) # for baseline model to get p value

  # model with and without foehn
  mod_modif     <- gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)
  model_baseline <- gnm(formula2, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # prediction with and without foehn
  pred_modif   <- crosspred(modif_rev, mod_modif, cen = 15.5)
  
    # p value extraction
  an <- lrtest(mod_modif, model_baseline)

  plot(pred_modif,              ## cumulative exposure
     "overall",
     ylab = "cumulative RR",
     xlab = "temperature [\u00B0C]",
     col = foehn_col,
     ci.arg = list(col = alpha(colour = foehn_col, .15)),
     lwd = 2,
     main ="",
     ylim = c(0.7,2.5),
     cex.axis = 0.7,
     cex.lab = 0.7)
  
  # p value in plot
  # text(-22, 1.5, labels = paste0("p-value = " , as.character(p_value)), pos = 4, cex = 0.7, )
  
  
  }

# dev.off()
```





```{r results_subgroups_interaction, echo=FALSE, fig.width=9, fig.height=9, fig.align='center'}
png("C:/Users/tinos/Documents/Master - Climate Science/3 - Master Thesis/plots/paper/rq2_subgroups_interaction_plot.png", width = 1800, height =1800, res = 300)

par(mfrow=c(3,3), 
    mar = c(3,3,.5,.5),
    mgp = c(1.8, .5, 0))

for (i in 1:length(groups_id[1:9])) {

  colvar  =  groups_id[i+1]

  formula1  <- as.formula(paste0(colvar, "~ cb.temp + modif_rev"))
  formula2 <- as.formula(paste0(colvar, "~ cb.temp"))
  
  
  # model with and without foehn
  mod_modif     <-gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)
  model_baseline <- gnm(formula2, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # prediction with and without foehn
  pred_modif   <- crosspred(modif_rev, mod_modif, cen = 15.5, by = 0.1)
  
  
  
      
  # p value extraction
  an <- anova(mod_modif, model_baseline)
  
  # chisq()
  deviance <- abs(an$Deviance[2])  # Deviance for Model 1 vs. Model 2
  df <- abs(an$Df[2])             # Degrees of freedom for Model 1 vs. Model 2
  p_value <- round(pchisq(deviance, df, lower.tail = FALSE), digits = 4)
                   
                   
                   
                   
  plot(pred_modif,              ## cumulative exposure
     "overall",
     ylab = "cumulative RR",
     xlab = "temperature [\u00B0C]",
     col = foehn_col,
     ci.arg = list(col = alpha(colour = foehn_col, .15)),
     lwd = 2,
     main ="",
     ylim = c(0.7,2.5),
     cex.axis = 0.7,
     cex.lab = 0.7)
  text(-14, 2.4, labels = paste0("(", letters[i], ")"), pos = 2)
  
  # p value in plot
  text(-14, 2, labels = paste0("p-value = " , as.character(p_value)), pos = 4)
  
  }

dev.off()
```

# below is not important

```{r results_subgroups_interaction, fig.width=9, fig.height=9, fig.align='center', warning=FALSE, message=FALSE}
par(mfrow=c(3,3), 
    mar = c(3,3,.5,.5),
    mgp = c(1.8, .5, 0))
dev.off()

# binary foehn
foehn_bin <- ifelse(data$f_id > 72, 0, 1)
foehn_bin_rev <- ifelse(foehn_bin == 1, 0, 1)

# modifier functions
modif <- cb.temp * foehn_bin
modif_rev <- cb.temp * foehn_bin_rev

# for (i in 1:length(groups_id[1:10])) {
for (i in 1:length(groups_id[2])) {

  colvar  =  groups_id[i]

  formula1  <- as.formula(paste0(colvar, "~ cb.temp + modif"))
  # formula2 <- as.formula(paste0(colvar, "~ cb.temp + modif_rev"))

  # model with and without foehn
  mod_modif     <- gnm(formula1, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)
  # mod_modif_rev <- gnm(formula2, data = data,  family=quasipoisson(), eliminate=stratum_dow, subset=ind_dow>0)

  # prediction with and without foehn
  pred_modif   <- crosspred(modif, mod_modif, cen = 15.5, by =0.1)
  # pred_modif_rev <- crosspred(modif_rev, mod_modif_rev, cen = 15.5, by = 0.1)
  # pred_modif   <- crosspred(cb.temp, mod_modif, cen = 15.5, by =0.1)
  # pred_modif_rev <- crosspred(cb.temp, mod_modif_rev, cen = 15.5, by = 0.1)
  
  plot(pred_modif,              ## cumulative exposure
     "overall",
     ylab = "cumulative RR",
     xlab = "temperature [\u00B0C]",
     col = foehn_col,
     ci.arg = list(col = alpha(colour = foehn_col, .2)),
     lwd = 2,
     main ="",
     # ylim = c(.99,1.02)
     )

  # print(colvar)
  print(pred_modif$allRRfit["24.7"])
  # print(pred_modif_rev$allRRfit["24.7"])
  # print(pred_modif$allRRlow["24.7"])
  # print(pred_modif$allRRhigh["24.7"])}
  
}



```



*Figure X: Subgroups interaction term*
 <br>

